---
title: "hakonRlab.github.io"
---

### **The Oslo model**

A lymph node invasion prediction model for prostate cancer patients

Sample size calculation was done using the pmsampsize package. None of the evaluated models have published the R^2^ so this was not included in our sample size calulation

```{r, eval=FALSE}
library(pmsampsize)
out <- pmsampsize(type = "b", cstatistic  = 0.8, parameters = 20, prevalence = 0.25, shrinkage = 0.8)
out$results_table

```

Imputing missing values with rmsb package

```{r, eval=FALSE}
library(rmsb)

set.seed(1999)         
mi <- aregImpute(~ pN + PSAutredn + ISUP3 + ISUP4 + ISUP5 + cT2a  + cT2b + cT2c + cT3 + PIRADS4 + PIRADS5  + cT_mr3a + cT_mr3b + MRindexTumorL1 + ProstVolum + BMI,
                 data=plndbayes_mrpNimp, n.impute=10, nk=3, pr=FALSE)

mi <- aregImpute(~ pN + PSAutredn + ISUP + PIRADS345  + mr_cT + MRindexTumorL1 + ProstVolum + dre_cT + BMI + PercPosCores, data=plndbayes_mrpNimp, n.impute=10, nk=4, pr=FALSE, boot.method = 'approximate bayesian')
```

Bayesian logistic regression model development followed many suggestions from the [Regression Modeling Strategies](https://hbiostat.org/rmsc/) by Frank Harrell. Used the rmsb package and the imputed data sets generated in the previous code chunk.

```{r, eval=FALSE}
akermodellenAlle <- stackMI(formula = pN ~ rcs(PSAutredn,4)*rcs(ProstVolum,4) + ISUP  + dre_cT  + PIRADS345  + mr_cT + rcs(MRindexTumorL1,4)  + rcs(BMI,4) +
                              rcs(PercPosCores,4), fitter = blrm, xtrans = mi, data=plndbayes_mrpNimp, iter = 4000, warmup = 2000 , chain = 4, file='akermodellenAlle.RDS')


plot(anova(akermodellenAlle)) # Relative explained variation of variables
```

Variable selection validation

```{r, eval=FALSE}
options(prType='html')

plot(anova(akermodellenAlle),  cex.lab = 1.2, cex.axis = 12)

plot(anova(akermodellen2),  cex.lab = 1.2, cex.axis = 12)

akaerteset <- vcov(akermodellenAlle, regcoef.only = TRUE, intercepts = "all")

akertest <- brms::get_refmodel.brmsfit(akermodellenAlle)


cvvs_imp1 <- cv_varsel(m3varsel,
                       validate_search = FALSE,
                       method = "forward",
                       ncluster_pred = 20,
                       nterms_max = 20,
                       verbose = FALSE)
plot(cvvs_imp1, stats = "mlpd", ranking_ntrems_max = NA)

plot(anova(akermodellen1))


val_aker_anov <- processMI(akermodellen, 'anova')
val_aker <- validate(akermodellenAlle, B=150, bw=TRUE, rule = 'p', sls = 0.1, type = 'individual')


cvvs_imp1 <- cv_varsel(m3varsel,
                       cv_method = "kfold",
                       method ="forward",
                       k=10,
                       ncluster_pred = 20,
                       nterms_max = 9,
                       parallel = TRUE,
                       verbose = FALSE)

plot(cvvs_imp1, stats = "mlpd", deltas  = TRUE)

suggest_size(cvvs_imp1)
```

The [Evaluating a logistic regression based prediction tool in R](https://darrendahly.github.io/post/homr) by Darren Dahly was used with minor modifications of the calibration plots.

```{r, eval=FALSE}

```

The resources at [Decission Curve Analyses](https://mskcc-epi-bio.github.io/decisioncurveanalysis/dca-tutorial.html) was used as a guideline for our net reduction analyses.
